---
title: "Fluffy_Bird"
output: html_notebook
---




#Library

Opening the libraries used in the coding. If the libraries does not work try the "install_packages" function. 

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

#Read data

The data is from an excel file and is read into R and named "Fluffu_data".

```{r}
# Read the Excel file
Fluffy_data <- read_excel("F_data.xlsx")
```







#Creating and looking at graphs (evolving)

First graph is just looking at the data. The only important graph is the Final Graph at the end of the coding (can be found at 136-170, or at the Outline overview "Final Graph"). The text and the values do not look nice, so i try to improve it to the next code/graph.

```{r}
ggplot(Fluffy_data, aes(x = Fluffiness, y = Temp_Area)) +
  geom_bar(stat = "identity", aes(fill = Fluffiness), color = "black") +
  labs(
    title = "Bird Fluffiness compared to temp",
    x = "Fluffiness Level",
    y = "Temperature in captured area"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "none"
  )
```


The text is still colliding and the coloring is very dark and not visualizing appealing. 

```{r}

ggplot(Fluffy_data, aes(x = Fluffiness, y = Temp_Area)) +
  geom_point(alpha = 0.5) +  # Use transparency for better visual overlap handling
  facet_grid(. ~ Fluffiness, scales = "free_x") +  # Facet by Fluffiness to compare per category
  labs(
    title = "Bird Fluffiness Compared to Area Temperature",
    y = "Temperature in Captured Area",
    x = "Fluffiness Level"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12)
  )

```

Text and size not appealing. 

```{r}

ggplot(Fluffy_data, aes(x = Fluffiness, y = Temp_Area)) +
  geom_jitter(aes(color = Fluffiness), width = 0.2, height = 0, alpha = 0.5) +  # Jitter the points horizontally
  facet_grid(. ~ Fluffiness, scales = "free_x") +  # Facet by Fluffiness to compare per category
  labs(
    title = "Bird Fluffiness Compared to Area Temperature",
    y = "Temperature in Captured Area",
    x = "Fluffiness Level"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12)
  )

```



Still some text colliding and i want the mean to show. 

```{r}

# Correctly convert Fluffiness to a factor with the desired order
Fluffy_data$Fluffiness <- factor(Fluffy_data$Fluffiness, 
                                  levels = c("Naked", "Light Fluff", "Medium Fluff", "Fluffy", "Cloud Fluffy", "Only Fluff"))

ggplot(Fluffy_data, aes(x = Fluffiness, y = Temp_Area)) +
  geom_jitter(aes(color = Fluffiness), width = 0.2, height = 0, alpha = 0.5) +
  facet_grid(. ~ Fluffiness, scales = "free_x") +
  labs(
    title = "Bird Fluffiness Compared to Area Temperature",
    y = "Temperature in Captured Area",
    x = "Fluffiness Level"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12),
    legend.position = "none" 
  )

```






#Final Graph
#Save as a png-image

The final version of the graphs. Satisfied with the visualization of the data. Using ggsave function to the the graph into a png-picture that can be used in a article (result part). 

```{r}

# Convert Fluffiness to a factor with the desired order
Fluffy_data$Fluffiness <- factor(Fluffy_data$Fluffiness, 
 levels = c("Naked", "Light Fluff", "Medium Fluff", "Fluffy"))

# Calculate mean temperatures for each Fluffiness level
mean_values <- Fluffy_data %>%
  group_by(Fluffiness) %>%
  summarize(mean_temp = mean(Temp_Area))


#save it a "Fluff_graph" so the graph-plot can be created into a png-picture
Fluff_graph <- ggplot(Fluffy_data, aes(x = Fluffiness, y = Temp_Area)) +
  geom_jitter(aes(color = Fluffiness), width = 0.2, height = 0, alpha = 0.6) + 
  # Add horizontal lines showing the mean temperature for each Fluffiness level
  geom_hline(data = mean_values, aes(yintercept = mean_temp), linetype = "dashed", color = "red") +
  facet_grid(. ~ Fluffiness, scales = "free_x") +  
   scale_y_continuous(breaks = seq(from = min(Fluffy_data$Temp_Area), 
                                  to = max(Fluffy_data$Temp_Area), by = 3)) +  # Customize breaks 
  labs(
    title = "Birds Fluffiness compared to Temperature in captured area",
    y = "Temperature in (C)",
    x = "Fluffiness Level"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 13, face = "bold"),
    strip.text = element_text(size = 12),
    legend.position = "none"
  )

#saving the graph as a png-picture
ggsave("fluffiness_temperature_plot.png", plot = Fluff_graph, width = 8, height = 6, dpi = 300)
```










